---
title: "class14"
author: "Xiao"
date: "2019/5/16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(BiocManager)
```


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE, row.names = 1)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
head(counts)
head(metadata)
```

```{r}
View(metadata)
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] ) / 4 
names(control.mean) <- counts$ensgene
```

lets check the corespondece of our two samples 
```{r}
all(colnames(counts) == metadata$id)
```

# Section 3. Walk through of gene expression analysis
Our metadata file tells us which columns of the countdata are drug treated and control
```{r}
control.inds <- metadata$dex == "control"
control <- metadata[control.inds, ]
control
head(counts[, control$id])
```

Determine the mean count values for all genes across control experiments
```{r}
control.mean <- rowSums(counts[, control$id]) / length(control$id)
```

Do the same things for treated experiments
```{r}
treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds, ]
treated
head(counts[, treated$id])
treated.mean <- rowSums(counts[, treated$id]) / length(treated$id)
```

Store the control.mean and treated.mean together for ease of use
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

What is the sum counts accross all genes in control and treated
```{r}
colSums(meancounts)
```

```{r}
plot(meancounts)
```

Plot a log-log sacle so we can see our data
```{r}
plot(meancounts, log = "xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
to.rm <- unique(which(meancounts[, 1:2] == 0, arr.ind = T)[, 1])
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes are up in the drug treated cells
```{r}
up.inds <- mycounts$log2fc > 2
sum(up.inds)
```

and how many are down in the drug treated cells
```{r}
down.inds <- mycounts$log2fc < -2
sum(down.inds)
```

```{r}
head(mycounts[up.inds,])
```

## Section 4. Adding annotation adata

We can add annotation from a supplied CSV file, such as those available from ENSEMBLE or UCSC. The annotables_grch38.csv annotation table links the unambiguous Ensembl gene ID to other useful annotation like the gene symbol, full gene name, location, Entrez gene ID, etc.

```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```
Use the **merge()** function to add the annotation data from the 'anno' object to out RNA-Seq results in 'mycounts'

```{r}
mycounts.anno <- merge(mycounts, anno, by.x = "row.names", by.y = "ensgene")
mycounts.anno
```

In cases where you don’t have a preferred annotation file at hand you can use other Bioconductor packages for annotation.
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

We will use the **mapIds()**function from bioconductor to add annotation data to our results
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                          keys = row.names(mycounts), # The gene names in out dataset
                          column = "SYMBOL",          # WHat new data do we want to add
                          keytype = "ENSEMBL",        # Format of our gene names
                          multiVals = "first")
```

```{r}
head(mycounts)
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                          keys = row.names(mycounts), # The gene names in out dataset
                          column = "ENTREZID",        # WHat new data do we want to add
                          keytype = "ENSEMBL",        # Format of our gene names 
                          multiVals = "first")
```

## Section 5. DESeq2 analysis
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
dds <- DESeqDataSetFromMatrix(countData = counts, 
                              colData = metadata, 
                              design = ~dex, 
                              tidy = T)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
summary(res)
```

```{r}
res05 <- results(dds, alpha = 0.05)
head(res05)
summary(res05)
```

```{r}
res01 <- as.data.frame(results(dds, alpha = 0.01))
head(res01)
```

Let's sort our order results by the adjested p-value
```{r}
ord.inds <- order( res01$padj )
write.csv(res01[ord.inds, ], file = "signal01_results.csv")
```

Volcano plot
```{r}
mycols <- rep("gray", nrow(res01))
mycols[res01$padj < 0.01] <- "black"
#mycols[abs(res01$log2FoldChange) > 2] <- "blue"
mycols[(res01$padj) < 0.01 & (abs(res01$log2FoldChange) > 2)] <- "red"
plot(res01$log2FoldChange, -log(res01$padj), ylab="-Log(P-value)", xlab="Log2(FoldChange)", col = mycols)

```

