---
title: 'Class 7: R functions ad packages'
author: "Xiao"
date: "2019/4/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# function revisited

We will source a file from online with our functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() fucntion

```{r}
rescale(1: 10)
```

Try the rescale2() functions that catches string inputs
```{r}
rescale2(c(1: 10, "string"))
```

# Find missing NA values in two vectors

Start with a simple example of he larger problem I am tring to solve

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
is.na(y)
```
Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```

Take the sum() to find out how many TRUE values we have and thus how many NAs we had in both x and y
```{r}
sum(is.na(x) & is.na(y))
```

Now I can make this into our first function...
```{r}
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, y)
```

Test, test, test
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```
```{r}
both_na(x, y2)
# gonna be 3 because the system will automatically recycle the first few elements in shorter vector the match their lengths.
```

Now lets try the new both_na2 function on our different length input vectors
```{r}
both_na2(x, y2)
```

```{r}
which(is.na(c(1, 2, NA, 4)))
```

```{r}
df1
```

```{r}
df2
```

Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs
```

```{r}
intersect(x, y)
```

```{r}
x %in% y
```

```{r}
# Code -> Extract function -> function name
gene_intersect <- function(x, y) {
  cbind( df1[ df1$IDs %in% df2$IDs, ],
         df2[ df2$IDs %in% df1$IDs, "exp"]  )
}
```

```{r}
gene_intersect(df1, df2)
```

```{r}
dropLow <- function(x) {
  i <- (sum(x) - min(x))/(length(x) - 1)
  message("The final average grade is ", i)
}
```
```{r}
x <- c(100, 100, 100, 100, 100, 100, 100, 90)
y <- c(100, 90, 90, 90, 90, 90, 97, 80)
dropLow(x)
dropLow(y)
```

